Practical Parametricity for GADTs

### 2020-09-02 20-gadts/inf-rules.tex
tweak condition for rel interp of Lans

\begin{dfn}\label{def:rel-transf}
A {\em $k$-ary $\oCPO$ relation transformer} $F$ is a triple $(F^1,
F^2,F^*)$, where $F^1,F^2 : [\oCPO^k,\oCPO]_{\omega_1}$ and $F^* :
[\oCPOR^k, \oCPOR]_{\omega_1}$ are such that if
$R_1:\oCPOR(A_1,B_1),...,R_k:\oCPOR(A_k,B_k)$ then $F^* \ol{R} :
\oCPOR(F^1 \ol{A}, F^2 \ol{B})$, and if $(\alpha_1, \beta_1) \in
\HomoCPOR(R_1,S_1),..., (\alpha_k, \beta_k) \in \HomoCPOR(R_k,S_k)$
then $F^* \ol{(\alpha, \beta)} = (F^1 \ol{\alpha}, F^2 \ol{\beta})$.
We define $F\ol{R}$ to be $F^*\overline{R}$ and
$F\overline{(\alpha,\beta)}$ to be $F^*\overline{(\alpha,\beta)}$.
\end{dfn}
The first condition of the first sentence of
Definition~\ref{def:rel-transf} entails that $F^*\ol R$ relates sups
of chains of pairwise related elements in $F^1 \ol A$ and $F^2 \ol
B$. The last condition of the first sentence of
Definition~\ref{def:rel-transf} expands to: if $\ol{(a,b) \in R}$
implies $\ol{(\alpha\,a,\beta\,b) \in S}$ then $(c,d) \in F^*\ol{R}$
implies $(F^1 \ol{\alpha}\,c,F^2 \ol{\beta}\,d) \in F^*\ol{S}$. When
convenient we identify a $0$-ary $\oCPO$ relation transformer
$(A,B,R)$ with $R : \oCPOR(A,B)$. We may also write $\pi_1 F$ for
$F^1$ and $\pi_2 F$ for $F^2$. We extend these conventions to $\oCPO$
relation environments, introduced in Definition~\ref{def:reln-env}
below, in the obvious way.

@@

\begin{dfn}\label{def:rel-sem}
The {\em relational interpretation}
$\oCPORsem{\cdot} : \F \to [\oCPORenv, \oCPOR]_{\omega_1}$ is defined by
\begin{align*}
  \oCPORsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPORsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPORsem{\Gamma; \emptyset \vdash \Nat^\Phi \,F\,G}\rho &= \{\eta
  : \lambda \ol{K}.\,\oCPORsem{\Gamma; \Phi \vdash
    F}\rho[\ol{\Phi := K}] \Rightarrow \lambda \ol{K}. \,\oCPORsem{
    \Gamma; \Phi \vdash G}\rho[\ol{\phi := K}]\}\\
  &=
  \{(\eta_1,\eta_2) \in \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G} (\pi_1 \rho) \times \oCPOsem{ 
    \Gamma;\emptyset
    \vdash \Nat^\Phi \,F\,G} (\pi_2
  \rho)~|~\\ 
  & \hspace{0.3in} \forall \ol{K = (K^1, K^2, K^*) : \oCPORT_{k}}.\\
  & \hspace{0.4in} ((\eta_1)_{\ol{K^1}},(\eta_2)_{\ol{K^2}}) \in
  (\oCPORsem{\Gamma; \Phi \vdash G}\rho[\ol{\Phi :=
      K}])^{\oCPORsem{\Gamma;\Phi\vdash F}\rho[\ol{\Phi := K}]} \}\\  
  \oCPORsem{\Gamma;\Phi \vdash \phi \ol{F}}\rho &=
  (\rho\phi)\ol{\oCPORsem{\Gamma;\Phi \vdash 
    F}\rho}\\
  \oCPORsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho +
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPORsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\  
   \oCPORsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho
  &= (\mu T_{H,\rho})\ol{\oCPORsem{\Gamma;\Phi,\ol\gamma \vdash
     G}\rho}\\
  \text{where }	T_{H,\rho}
    &= (T^\set_{H,\pi_1\rho}, T^\set_{H,\pi_2\rho}, T^\rel_{H,\rho}) \\
  \text{and } T^\oCPOR_{H,\rho}\,F
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\rho[\phi :=
    F][\ol{\alpha := R}]\\
  \text{and } T^\oCPOR_{H,\rho}\,\delta
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\id_\rho[\phi :=
    \delta][\ol{\alpha := \id_{\ol{R}}}] \\
  \oCPORsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
    F)\ol{A}}\rho &= \{ (t_1, t_2) : \oCPOsem{\Gamma;\Phi \vdash
    (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\pi_1 \rho \,\times\,
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
    F)\ol{A}}\pi_2 \rho\\
  &\hspace{0.3in}|~{\color{blue}
    ((\eta_1)_{\ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_1 \rho}} t_1,
(\eta_2)_{\ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_2 \rho}} t_2) \in}\\
  &\hspace{0.3in}{\color{blue}
(\textit{Lan}_{\ol{\oCPORsem{\Gamma;
          \ol\alpha \vdash K}\rho[\ol{\alpha := \Eq_{\_}}]}}
    \oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\ol{\alpha :=
        \Eq_{\_}}])
    \ol{\oCPORsem{\Gamma; \Phi \vdash A}\rho}} \}  
  %\forall \sigma : \setenv, (
  %(\eta^\sigma_1)_{\ol{\oCPOsem{\Gamma;\Phi \vdash A}\sigma }} t_1,
  %(\eta^\sigma_2)_{\ol{\oCPOsem{\Gamma;\Phi \vdash A}\sigma }} t_2) \in\\
  %&\hspace{0.5in} (\textit{Lan}_{\ol{\oCPORsem{\Gamma; \ol\alpha
  %      \vdash K}\Eq_\sigma[\ol{\alpha := \Eq_{\_}}]}}
  %\oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_\sigma[\ol{\alpha :=
  %    \Eq_{\_}}]) \ol{\oCPORsem{\Gamma; \Phi \vdash A}\Eq_\sigma} \}\\
\end{align*}
\end{dfn}
{\color{blue} where $\eta_i$ is for the (mixed $\set/\rel$) diagram
  top p2 of Daniel's notes for relational environment $\Eq_{\pi_i
    \rho}[\alpha := \Eq_{\_}]$.}

>>>>

\begin{dfn}\label{def:rel-transf}
A {\em $k$-ary $\oCPO$ relation transformer} $F$ is a triple $(F^1,
F^2,F^*)$, where $F^1,F^2 : [\oCPO^k,\oCPO]_{\omega_1}$ and $F^* :
[\oCPOR^k, \oCPOR]_{\omega_1}$ are such that if
$R_1:\oCPOR(A_1,B_1),...,R_k:\oCPOR(A_k,B_k)$ then $F^* \ol{R} :
\oCPOR(F^1 \ol{A}, F^2 \ol{B})$, and if $(\alpha_1, \beta_1) \in
\HomoCPOR(R_1,S_1),..., (\alpha_k, \beta_k) \in \HomoCPOR(R_k,S_k)$
then $F^* \ol{(\alpha, \beta)} = (F^1 \ol{\alpha}, F^2 \ol{\beta})$.
We define $F\ol{R}$ to be $F^*\overline{R}$ and
$F\overline{(\alpha,\beta)}$ to be $F^*\overline{(\alpha,\beta)}$.
\end{dfn}
The first condition of the first sentence of
Definition~\ref{def:rel-transf} entails that $F^*\ol R$ relates sups
of chains of pairwise related elements in $F^1 \ol A$ and $F^2 \ol
B$. The last condition of the first sentence of
Definition~\ref{def:rel-transf} expands to: if $\ol{(a,b) \in R}$
implies $\ol{(\alpha\,a,\beta\,b) \in S}$ then $(c,d) \in F^*\ol{R}$
implies $(F^1 \ol{\alpha}\,c,F^2 \ol{\beta}\,d) \in F^*\ol{S}$. When
convenient we identify a $0$-ary $\oCPO$ relation transformer
$(A,B,R)$ with $R : \oCPOR(A,B)$. We may also write $\pi_1 F$ for
$F^1$ and $\pi_2 F$ for $F^2$. {\color{blue} Without loss of
  generality, we may assume that $\pi_1$ and $\pi_2$ are surjective on
  relations.} We extend these conventions to $\oCPO$ relation
environments, introduced in Definition~\ref{def:reln-env} below, in
the obvious way.

@@

\begin{dfn}\label{def:rel-sem}
The {\em relational interpretation}
$\oCPORsem{\cdot} : \F \to [\oCPORenv, \oCPOR]_{\omega_1}$ is defined by
\begin{align*}
  \oCPORsem{\Gamma;\Phi \vdash \zerot}\rho &= 0\\
  \oCPORsem{\Gamma;\Phi \vdash \onet}\rho &= 1\\
  \oCPORsem{\Gamma; \emptyset \vdash \Nat^\Phi \,F\,G}\rho &= \{\eta
  : \lambda \ol{K}.\,\oCPORsem{\Gamma; \Phi \vdash
    F}\rho[\ol{\Phi := K}] \Rightarrow \lambda \ol{K}. \,\oCPORsem{
    \Gamma; \Phi \vdash G}\rho[\ol{\phi := K}]\}\\
  &=
  \{(\eta_1,\eta_2) \in \oCPOsem{\Gamma; \emptyset
    \vdash \Nat^\Phi
    \,F\,G} (\pi_1 \rho) \times \oCPOsem{ 
    \Gamma;\emptyset
    \vdash \Nat^\Phi \,F\,G} (\pi_2
  \rho)~|~\\ 
  & \hspace{0.3in} \forall \ol{K = (K^1, K^2, K^*) : \oCPORT_{k}}.\\
  & \hspace{0.4in} ((\eta_1)_{\ol{K^1}},(\eta_2)_{\ol{K^2}}) \in
  (\oCPORsem{\Gamma; \Phi \vdash G}\rho[\ol{\Phi :=
      K}])^{\oCPORsem{\Gamma;\Phi\vdash F}\rho[\ol{\Phi := K}]} \}\\  
  \oCPORsem{\Gamma;\Phi \vdash \phi \ol{F}}\rho &=
  (\rho\phi)\ol{\oCPORsem{\Gamma;\Phi \vdash 
    F}\rho}\\
  \oCPORsem{\Gamma;\Phi \vdash F+G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho +
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\
  \oCPORsem{\Gamma;\Phi \vdash F\times G}\rho &=
  \oCPORsem{\Gamma;\Phi \vdash F}\rho \times
  \oCPORsem{\Gamma;\Phi \vdash G}\rho\\  
   \oCPORsem{\Gamma;\Phi,\ol\gamma \vdash (\mu \phi.\lambda
    \ol{\alpha}. H)\ol{G}}\rho
  &= (\mu T_{H,\rho})\ol{\oCPORsem{\Gamma;\Phi,\ol\gamma \vdash
     G}\rho}\\
  \text{where }	T_{H,\rho}
    &= (T^\set_{H,\pi_1\rho}, T^\set_{H,\pi_2\rho}, T^\rel_{H,\rho}) \\
  \text{and } T^\oCPOR_{H,\rho}\,F
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\rho[\phi :=
    F][\ol{\alpha := R}]\\
  \text{and } T^\oCPOR_{H,\rho}\,\delta
    &= \lambda \ol{R}. \oCPORsem{
      \Gamma;\ol\gamma,\phi,\ol{\alpha} \vdash H}\id_\rho[\phi :=
    \delta][\ol{\alpha := \id_{\ol{R}}}] \\
  \oCPORsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
    F)\ol{A}}\rho &= \{ (t_1, t_2) : \oCPOsem{\Gamma;\Phi \vdash
    (\Lan^{\ol{\alpha}}_{\ol{K}} F)\ol{A}}\pi_1 \rho \,\times\,
  \oCPOsem{\Gamma;\Phi \vdash (\Lan^{\ol{\alpha}}_{\ol{K}}
    F)\ol{A}}\pi_2 \rho\\
  &\hspace{0.3in}|~{\color{blue}
    (\nu_1 ((\eta_1)_{\ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_1 \rho}} t_1),
\nu_2 ((\eta_2)_{\ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_2 \rho}} t_2)) \in}\\
  &\hspace{0.3in}{\color{blue}
(\textit{Lan}_{\ol{\oCPORsem{\Gamma;
          \ol\alpha \vdash K}\rho[\ol{\alpha := \Eq_{\_}}]}}
    \oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\rho[\ol{\alpha :=
        \Eq_{\_}}])
    \ol{\oCPORsem{\Gamma; \Phi \vdash A}\rho}} \}  
  %\forall \sigma : \setenv, (
  %(\eta^\sigma_1)_{\ol{\oCPOsem{\Gamma;\Phi \vdash A}\sigma }} t_1,
  %(\eta^\sigma_2)_{\ol{\oCPOsem{\Gamma;\Phi \vdash A}\sigma }} t_2) \in\\
  %&\hspace{0.5in} (\textit{Lan}_{\ol{\oCPORsem{\Gamma; \ol\alpha
  %      \vdash K}\Eq_\sigma[\ol{\alpha := \Eq_{\_}}]}}
  %\oCPORsem{\Gamma; \Phi, \ol\alpha \vdash F}\Eq_\sigma[\ol{\alpha :=
  %    \Eq_{\_}}]) \ol{\oCPORsem{\Gamma; \Phi \vdash A}\Eq_\sigma} \}\\
\end{align*}
\end{dfn}
{\color{blue} where $\eta_i$ is for the (mixed $\set/\rel$) diagram
  top p2 of Daniel's notes for relational environment $\Eq_{\pi_i
    \rho}[\alpha := \Eq_{\_}]$ and $\nu_i$ is the unique morphism in
  $\set$ from \[(\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
        K}\pi_i\rho[\ol{\alpha := \_}]}} \oCPOsem{\Gamma; \Phi,
    \ol\alpha \vdash F}\pi_i\rho[\ol{\alpha := \_}])
  \ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_i\rho}\] to
\[\pi_i(\textit{Lan}_{\ol{\oCPOsem{\Gamma; \ol\alpha \vdash
      K}\rho[\ol{\alpha := \Eq_{\_}}]}} \oCPOsem{\Gamma; \Phi,
  \ol\alpha \vdash F}\rho[\ol{\alpha := \Eq_{\_}}])
\ol{\oCPOsem{\Gamma; \Phi \vdash A}\pi_i\rho}\]
given by the colimit formulation of left Kan extensions.}
